Welch-Lynch clock synchronization protocol \\\textit{Modelling in UPPAAL}

### 2013-04-15 paper/main.tex
Crappy first attempt at explaining optimizations (5/7)

>>>>

In 2004, a model was made that models the protocol in almost its most general form. However, fault-tolerance is completely ignored and a trick is applied to model the floating point numbers that correspond to clocks. This has been our starting point of reference. The most important automaton, the description of a process, can be seen in figure~\ref{fig:original_process}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{original_process}
\label{fig:original_process}
\end{figure}

It was soon getting clear that all other optimizations had to be applied as well and even additional ones. Below is a complete list of all optimization steps that have been used and how they contribute to making the model verifiable.

\begin{enumerate}
\item \textbf{C-like functions} \\
	The \UPPAAL C-like declaration language has been enriched extensively over the years. For instance, in 2004, one was not able to use functions and complicated constructions had to be thought off to simulate the behaviour. In figure~\ref{fig:original_process}, the right part uses a committed state to iterate over an array to determine its minimal and maximal values. A similar structure is seen in the channel automaton. Luckily the new syntax allows us to replace these structures by C-like functions. Due to the nature of a committed state, this does not directly highly decrease the state space, but it allows for the removal of some variables, transitions and complex update statements. This at least makes the model more comprehensible. The result can be seen in figure~\ref{fig:step1_process}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{step1_process}
\label{fig:step1_process}
\end{figure}

\item \textbf{Select statements} \\
	The 2004 model uses two \textit{choice variable} automatons to let each process and each channel non-deterministically choose which message to receive. The automatons are used to iterate over all process instances. Such choice variables can now easily be replaced by select statements that range over some type \texttt{id\_t}. This reduces the state space. All iterating is now done implicitly within one single transition, without branching the search space.

\item \textbf{Rewriting variables} \\
	The \UPPAAL verification engine can not deal with unbounded integers very well. To solve this, it was proposed in 2004 to rewrite lots of variables in such a way that one can work with bounded variables. The main idea is to introduce the notion of a round explicitly in the description of a process. This is implemented by subtracting the round index multiplied with the time between rounds from all timer variables. Now effectively, these timers can only take values up to the time between rounds, a constant \(P\). Then some invariants are used to further reduce the number of variables. The correctness property is also altered by this, but it can be proven that the systems are equivalent.

\item \textbf{Specific three process instance} \\
	In 2004, it was probably noticed that it turned out to be hard to verify instances of the protocol using \UPPAAL. Therefore, the researchers tried to restrict themselves to three processes. This allows for a more direct computation of the minimum and maximum values in the right part of the process automaton. The variables \texttt{ARR1}, \texttt{min1} and \texttt{max1} are replaced by a single variable \texttt{SUM}. Removing variables implies reducing the search space.

\item \textbf{Time jumps} \\
	When traces were simulated within the model, it was noticed that for the larger part of the time, most variables were not changing values, except for the \texttt{local\_time} variable which was simply incrementing. When the counter reached 0, all process instances started to send out and receive messages. Before and after that, nothing changed within the system. This lead to the invention of a time jump optimization, where a process is allowed to simply skip the parts where nothing interesting is happening. Of course it has been made sure that the time jumps do not influence the total behaviour of the system in any way. This optimization has been shown to greatly reduce the state space. The resulting process automaton can be seen in figure~\ref{fig:step3minimized_process}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{step3minimized_process}
\label{fig:step3minimized_process}
\end{figure}

\item \textbf{Symmetry reduction} \\
%TODO

\item \textbf{Extended guards} \\
%TODO
\end{enumerate}
